Scattering from compact objects: \\ Regge poles and the Complex Angular Momentum method

### 2020-04-14 rainbowCAM.tex
Version submited to arXiv
corrected typo in eq 50 (a) - imaginary background integral
tom: reformatted eq 50 (c) after correcting + added files pertaining to the referees comments and our response. Now need to collate comments for a reply and to implement any changes
made some changes for referee comments - see summary of changes

%\documentclass[aps,prd,longbibliography,reprint,twocolumn,amsmath,amssymb,amsfonts,showpacs,superscriptaddress]{revtex4-1}%reprint
\documentclass[aps,prd,longbibliography,reprint,twocolumn,amsmath,amssymb,amsfonts,showpacs,footnote,superscriptaddress]{revtex4-1}%reprint

@@

Here we follow the approach of Zhang, Wu and Leung~\cite{Zhang:2011pq}, who applied the WKB method to determine the axial \textit{w}-modes of a variety of stellar models. We adapt their method to obtain analytical approximations for the ``broad resonances'' for massless waves on a stellar background. The starting point is the radial equation (\ref{H_Radial_equation}) with either the effective potential for the scalar field (\ref{Inside_Potentiel}), or for axial gravitational perturbations (\ref{RW_Potentiel_axialGW}).

@@

In the interior of the star, the radial function $f(r)$ is
\begin{equation}
f(r) = f_0 \left(1 + O\left(\frac{M r^2}{R^3} \right) \right),
\end{equation}
where $f_0$ is a constant, and $h(r) = 1 - 2Mr^2/R^3$. In the vicinity of $r=0$, we may solve Eq.~(\ref{eq:tortoise}) to obtain the tortoise coordinate $r_\ast(r)$, or inversely,
\begin{equation}
r = \sqrt{f_0} \, r_\ast \left(1 + O\left(\frac{M r_\ast^2}{R^3} \right) \right) . \label{eq:rrstar}
\end{equation}
Now let us consider the radial equation (\ref{H_Radial_equation}) in the high-frequency regime with $\ell \gg 1$, such that we may neglect all but the angular momentum terms in the effective potential (\ref{Inside_Potentiel}). Inserting (\ref{eq:rrstar}) and neglecting the quadratic corrections leads to a comparison equation
\begin{equation}
\left[\frac{d^{2}}{dr_{\ast}^{2}}+\omega^{2} - \frac{\ell (\ell+1)}{r_\ast^2} \right] \phi_{\omega\ell}= 0,
\end{equation}
with the regular interior solution
\begin{equation}\label{Approx_origin}
  \phireg \underset{\omega \to \infty}{=} \omega r_* j_{\lambda-1/2} (\omega r_*),
\end{equation}
where $j_{\lambda-1/2}(\cdot)$ is the spherical Bessel function of the first kind.

@@

@@

Now, we ``deform'' the contour ${\cal C}$ in Eq.~(\ref{SW_Scalar_Scattering_amp}) in order to collect, by using Cauchy's theorem, the Regge poles contributions. This is achieved by following, \textit{mutatis mutandis}, the approach developed in Ref~\cite{Folacci:2019cmc} (see more particularly Sec. IIB~3 and Fig.~1). We obtain
%
\begin{equation}\label{CAM_Scalar_Scattering_amp_tot}
\hat{f} (\omega, \theta) =  \hat{f}^\text{\tiny{B}} (\omega, \theta) +  \hat{f}^\text{\tiny{RP}} (\omega, \theta)
\end{equation}
where
\begin{subequations}\label{CAM_Scalar_Scattering_amp_decomp}
\begin{equation}\label{CAM_Scalar_Scattering_amp_decomp_Background}
\hat{f}^\text{\tiny{B}} (\omega, \theta) = \hat{f}^\text{\tiny{B},\tiny{Re}} (\omega, \theta)+ \hat{f}^\text{\tiny{B},\tiny{Im}} (\omega, \theta)
\end{equation}
is a background integral contribution with
\begin{equation}\label{CAM_Scalar_Scattering_amp_decomp_Background_a}
\hat{f}^\text{\tiny{B},\tiny{Re}} (\omega, \theta) = \frac{1}{\pi \omega} \int_{{\cal C}_{-}} d\lambda \, \lambda S_{\lambda -1/2}(\omega) Q_{\lambda -1/2}(\cos \theta +i0)
\end{equation}
and
\begin{eqnarray}\label{CAM_Scalar_Scattering_amp_decomp_Background_b}
\hat{f}^\text{\tiny{B},\tiny{Im}} && (\omega, \theta) = \frac{1}{2 \omega}\left(\int_{+i\infty}^{0} d\lambda \, \left[S_{\lambda -1/2}(\omega) P_{\lambda-1/2} (-\cos \theta) \right. \right.\nonumber \\
&& -\left. \left. S_{-\lambda -1/2}(\omega) e^{i \pi \left(\lambda+1/2\right)}P_{\lambda-1/2} (\cos \theta) \right]\lambda\right).
\end{eqnarray}
\end{subequations}
The second term in Eq.~(\ref{CAM_Scalar_Scattering_amp_tot}) ,
\begin{eqnarray}\label{CAM_Scalar_Scattering_amp_decomp_RP}
& & \hat{f}^\text{\tiny{RP}} (\omega, \theta) = -\frac{i \pi}{\omega}    \sum_{n=1}^{+\infty}   \frac{ \lambda_n(\omega) r_n(\omega)}{\cos[\pi \lambda_n(\omega)]}  \nonumber \\
&&  \qquad\qquad \qquad\qquad \times  P_{\lambda_n(\omega) -1/2} (-\cos \theta),
\end{eqnarray}
is a sum over the Regge poles lying in the first quadrant of the CAM plane. Of course, Eqs.~(\ref{CAM_Scalar_Scattering_amp_tot}), (\ref{CAM_Scalar_Scattering_amp_decomp}) and (\ref{CAM_Scalar_Scattering_amp_decomp_RP}) provide an exact representation of the scattering amplitude $\hat{f} (\omega, \theta)$ for the scalar field, equivalent to the initial partial wave expansion (\ref{Scalar_Scattering_amp}). From this CAM representation, we can extract the contribution $\hat{f}^\text{\tiny{RP}} (\omega, \theta)$ given by (\ref{CAM_Scalar_Scattering_amp_decomp_RP}) which, as a sum over Regge poles, is only an approximation of $\hat{f} (\omega, \theta)$, and which provides us with a corresponding approximation of the differential scattering cross section via (\ref{Scalar_Scattering_diff}).

>>>>

\documentclass[aps,prd,longbibliography,reprint,twocolumn,amsmath,amssymb,amsfonts,showpacs,footnote,superscriptaddress]{revtex4-1}%reprint

@@

Here we follow the approach of Zhang, Wu and Leung~\cite{Zhang:2011pq}, who applied the WKB method to determine the axial \textit{w}-modes of a variety of stellar models. We adapt their method to obtain analytical approximations for the ``broad resonances'' for massless waves on a stellar background. The starting point is the radial equation (\ref{H_Radial_equation}) with either the effective potential for the scalar field (\ref{Inside_Potentiel}), or for axial gravitational perturbations (\ref{RW_Potentiel_axialGW}). 
It is valid only for models with $R/M>3$.

@@

In the interior of the star, the radial function $f(r)$ is
\begin{equation}\label{eq:f_exp}
f(r) = f_0 \left(1 + O\left(\frac{M r^2}{R^3} \right) \right),
\end{equation}
where $f_0$ is a constant, and $h(r) = 1 - 2Mr^2/R^3$. In the vicinity of $r=0$, we may solve Eq.~(\ref{eq:tortoise}) to obtain the tortoise coordinate $r_\ast(r)$, or inversely,
\begin{equation}
r = \sqrt{f_0} \, r_\ast \left(1 + O\left(\frac{M r_\ast^2}{R^3} \right) \right) . \label{eq:rrstar}
\end{equation}
Now let us consider the radial equation (\ref{H_Radial_equation}) in the high-frequency regime with $\ell \gg 1$, such that we may neglect all but the angular momentum terms in the effective potential (\ref{Inside_Potentiel}). Inserting (\ref{eq:rrstar}) and neglecting the quadratic corrections leads to a comparison equation
\begin{equation}
\left[\frac{d^{2}}{dr_{\ast}^{2}}+\omega^{2} - \frac{\ell (\ell+1)}{r_\ast^2} \right] \phi_{\omega\ell}= 0,
\end{equation}
with the regular interior solution
\begin{equation}\label{Approx_origin}
  \phireg \underset{\omega \to \infty}{=} \omega r_* j_{\lambda-1/2} (\omega r_*),
\end{equation}
where $j_{\lambda-1/2}(\cdot)$ is the spherical Bessel function of the first kind.

@@

As $M/R \rightarrow 0$ (a large dilute star), the poles move closer to the real axis. When $M/R=0$ (i.e. no star), the potential is $C^{\infty}$ and the branch does not exist.

For any stellar model (such as a polytropic fluid sphere), the potential may in general be $C^{N}$ at the surface (and $C^{\infty}$ elsewhere). In this case, $\mathcal{R}$ and thus the broad resonance branch depends on the discontinuity in the $N$-th derivative of the potential at $r=R$. We do not go into detail here, but see Eq.~4.5 of Ref. \cite{Zhang:2011pq}.

@@

Now, we ``deform'' the contour ${\cal C}$ in Eq.~(\ref{SW_Scalar_Scattering_amp}) in order to collect, by using Cauchy's theorem, the Regge poles contributions. This is achieved by following, \textit{mutatis mutandis}, the approach developed in Ref~\cite{Folacci:2019cmc} (see more particularly Sec. IIB~3 and Fig.~1). We obtain
%
\begin{equation}\label{CAM_Scalar_Scattering_amp_tot}
\hat{f} (\omega, \theta) =  \hat{f}^\text{\tiny{B}} (\omega, \theta) +  \hat{f}^\text{\tiny{RP}} (\omega, \theta)
\end{equation}
where
\begin{subequations}\label{CAM_Scalar_Scattering_amp_decomp}
\begin{equation}\label{CAM_Scalar_Scattering_amp_decomp_Background}
\hat{f}^\text{\tiny{B}} (\omega, \theta) = \hat{f}^\text{\tiny{B},\tiny{Re}} (\omega, \theta)+ \hat{f}^\text{\tiny{B},\tiny{Im}} (\omega, \theta)
\end{equation}
is a background integral contribution with
\begin{equation}\label{CAM_Scalar_Scattering_amp_decomp_Background_a}
\hat{f}^\text{\tiny{B},\tiny{Re}} (\omega, \theta) = \frac{1}{\pi \omega} \int_{{\cal C}_{-}} d\lambda \, \lambda S_{\lambda -1/2}(\omega) Q_{\lambda -1/2}(\cos \theta +i0)
\end{equation}
and
\begin{eqnarray}\label{CAM_Scalar_Scattering_amp_decomp_Background_b}
\hat{f} && ^\text{\tiny{B},\tiny{Im}}  (\omega, \theta) = \frac{1}{2 \omega}\left(\int_{+i\infty}^{0} d\lambda \, \left[S_{\lambda -1/2}(\omega) P_{\lambda-1/2} (-\cos \theta) \right. \right.\nonumber \\
 - && \left.  \left. S_{-\lambda -1/2}(\omega) e^{i \pi \left(\lambda+1/2\right)}P_{\lambda-1/2} (\cos \theta) \right] \frac{\lambda}{\cos (\pi \lambda) } \right).
\end{eqnarray}
\end{subequations}
The second term in Eq.~(\ref{CAM_Scalar_Scattering_amp_tot}) ,
\begin{eqnarray}\label{CAM_Scalar_Scattering_amp_decomp_RP}
& & \hat{f}^\text{\tiny{RP}} (\omega, \theta) = -\frac{i \pi}{\omega}    \sum_{n=1}^{+\infty}   \frac{ \lambda_n(\omega) r_n(\omega)}{\cos[\pi \lambda_n(\omega)]}  \nonumber \\
&&  \qquad\qquad \qquad\qquad \times  P_{\lambda_n(\omega) -1/2} (-\cos \theta),
\end{eqnarray}
is a sum over the Regge poles lying in the first quadrant of the CAM plane. Of course, Eqs.~(\ref{CAM_Scalar_Scattering_amp_tot}), (\ref{CAM_Scalar_Scattering_amp_decomp}) and (\ref{CAM_Scalar_Scattering_amp_decomp_RP}) provide an exact representation of the scattering amplitude $\hat{f} (\omega, \theta)$ for the scalar field, equivalent to the initial partial wave expansion (\ref{Scalar_Scattering_amp}). From this CAM representation, we can extract the contribution $\hat{f}^\text{\tiny{RP}} (\omega, \theta)$ given by (\ref{CAM_Scalar_Scattering_amp_decomp_RP}) which, as a sum over Regge poles, is only an approximation of $\hat{f} (\omega, \theta)$, and which provides us with a corresponding approximation of the differential scattering cross section via (\ref{Scalar_Scattering_diff}).
